A User-Centric Approach to the Design and Consequences of Recommender Systems

### 2019-11-19 paper/filter_bubbles.tex
update figures'

\documentclass[sigconf, review, anonymous]{acmart}
\settopmatter{printacmref=false}
\usepackage[T1]{fontenc}
%\usepackage{lmodern}
\usepackage{float}
\usepackage{enumitem}
\newtheorem{finding}{Finding}
%
% defining the \BibTeX command - from Oren Patashnik's original BibTeX documentation.
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08emT\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

@@

%
% The "title" command has an optional parameter, allowing the author to define a "short title" to be used in page headers.
\title{A User-Centric Approach to the Design and Consequences of Recommender Systems}

%
% By default, the full list of authors will be used in the page headers. Often, this list is too long, and will overlap
% other information printed in the page headers. This command allows the author to define a more concise list
% of authors' names for this purpose.

@@

We simulate over risk-aversion parameters $\gamma \in \{ 0, 0.3, 0.6, 1, 5 \}$, standard-deviation $\sigma \in \{ 0.25, 0.5, 1.0, 2.0, 4.0 \}$, correlation parameters $\rho\in \{ 0, 0.1, 0.3, 0.5, 0.7, 0.9 \} $ and degree of idiosyncrasy of individual valuations $\beta \in \{ 0, 0.4, 0.8, 1, 2, 5\}$. When we consider results varying a single parameter, we group the results over the other parameters and provide reports varying only the parameter of interest. We report results for a relatively small consumption history $T=20$ with a product space size $N=200$. Unless otherwise reported, our results are robust to varying $N$.

@@

\begin{figure}[t]
\includegraphics[width=0.9\linewidth]{figures/gamma_consumption_dist_N_200T_20.pdf}\\
\includegraphics[width=0.9\linewidth]{figures/gamma_consumption_dist_N_200T_20_partial.pdf}\\
\includegraphics[width=0.9\linewidth]{figures/gamma_consumption_dist_N_200T_20_omni.pdf}\\
\caption{Effect of Risk Aversion on Local Consumption for No Recommendation (Top), Recommendation (Middle) and Oracle (Bottom)}
\label{fig:no_rec_risk_aversion}
\end{figure}

@@

\begin{figure}[ht]
\includegraphics[width=.85\linewidth]{figures/gamma_diversity_N_200_T_20}
\captionof{figure}{Risk Aversion and Diversity}\label{fig:risk_aversion_diversity}
\end{figure}

\begin{finding}\label{finding_welfare_gap}
The welfare gap between the no recommendation and the recommendation regimes and that between the no recommendation and the oracle regimes are decreasing in $\rho$.
\end{finding}
However, interestingly, the decrease in diversity is \textit{not} associated with a decline in welfare. In fact, welfare stays roughly the same in the oracle and recommendation cases. The welfare gap between the no recommendation case and the other two cases \textit{decreases} as the diversity gap increases.
\par
\begin{figure}[t]
\includegraphics[width=.85\linewidth]{figures/rho_diversity_N_200_T_20.pdf}\\
\includegraphics[width=.85\linewidth]{figures/rho_welfare_N_200_T_20.pdf}\\
\captionof{figure}{Welfare, Diversity, and Correlation}\label{fig:diversity_correlation}
\end{figure}

@@

This only happens since $\gamma = 0$ leads to users only caring about the expected value of the product. However, as we saw in Findings \ref{finding_local_consumption} and \ref{finding_diversity}, increasing $\gamma$ can lead to lower diversity and increasingly local consumption due to the fact that the degree of uncertainty now impacts users' choices. This weakens the negative relationship between diversity and welfare since both negative and positive experiences with a good reduce uncertainty about surrounding goods. This leads to the inverted-U shape found in Figure \ref{fig:diversity_welfare_ra} when $\gamma$ is relatively large (e.g $\gamma = 5$) though diversity and welfare are virtually uncorrelated in the data. In the recommendation and oracle regimes, under risk neutrality ($\gamma=0$), welfare and diversity and uncorrelated, while under risk aversion ($\gamma=5$), it is possible to observe in Figure \ref{fig:diversity_welfare_ra_partial} an actual positive relation between diversity and welfare as recommendations are able to reduce uncertainty and facilitate exploration of the product space.
\par
\begin{figure}[t]
\includegraphics[width=1.05\linewidth]{figures/diversity_welfare_rn.pdf}\\
\includegraphics[width=1.05\linewidth]{figures/diversity_welfare_ra.pdf}\\
\captionof{figure}{Diversity vs. Welfare, No-Recommendation, $\gamma = 0$ (Top), Diversity vs. Welfare, $\gamma = 5$ (Bottom)}\label{fig:diversity_welfare_ra}
\end{figure}
\begin{figure}[t]
\includegraphics[width=1.0\linewidth]{figures/diversity_welfare_rn_partial.pdf}\\
\includegraphics[width=1.0\linewidth]{figures/diversity_welfare_ra_partial.pdf}\\
\captionof{figure}{Diversity vs. Welfare, Recommendation, $\gamma = 0$ (Top), Diversity vs. Welfare, $\gamma = 5$ (Bottom)}\label{fig:diversity_welfare_ra_partial}
\end{figure}

@@

\begin{figure}[t]
\includegraphics[width=.7\linewidth]{figures/beta_homogeneity_N_200_T_20}
\captionof{figure}{Strength of Recommendation and Homogeneity}\label{fig:beta_homo}
\end{figure}
~

\begin{figure}[t]
\includegraphics[width=.7\linewidth]{figures/rho_homogeneity_N_200_T_20}
\captionof{figure}{Correlation and Homogeneity}\label{fig:cor_homo}
\end{figure}

@@

As a result, several alternative recommendation evaluation metrics, such as serendipity \cite{kotkov2016survey}, coverage \cite{ge2010beyond}, novelty \cite{vargas2011rank}, and many others, have been proposed in order to develop recommender systems that produce recommendations that are more useful for users. We use our model to motivate an alternative recommendation evaluation metric that provides an alternative perspective on the serendipity metric proposed in the literature. Serendipitous recommendations are said to ``have the quality of being both unexpected and useful" \cite{maksai2015predicting} and we use this as a guide towards our definition.

@@

We have used the insights from our model to guide the evaluation and design of recommender systems. The key insight is that a critical component of designing good recommendations is understanding what user behavior would be without recommendation and then understanding how recommendation would change their consumption. Operationalizing this requires collecting information on user beliefs about the goods in the product space, which is not data that is traditionally collected for recommender systems.
\par

>>>>

\documentclass[sigconf]{acmart}
\settopmatter{printacmref=false}
\usepackage[T1]{fontenc}
%\usepackage{lmodern}
\usepackage{float}
\usepackage{enumitem}
\newtheorem{finding}{Finding}
%
% defining the \BibTeX command - from Oren Patashnik's original BibTeX documentation.
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08emT\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

@@

%
% The "title" command has an optional parameter, allowing the author to define a "short title" to be used in page headers.
\title{A User-Centric Approach to the Design and Consequences of Recommender Systems}
%
% By default, the full list of authors will be used in the page headers. Often, this list is too long, and will overlap
% other information printed in the page headers. This command allows the author to define a more concise list
% of authors' names for this purpose.

@@

We simulate over risk-aversion parameters $\gamma \in \{ 0, 0.3, 0.6, 1, 5 \}$, standard-deviation $\sigma \in \{ 0.25, 0.5, 1.0, 2.0, 4.0 \}$, correlation parameters $\rho\in \{ 0, 0.1, 0.3, 0.5, 0.7, 0.9 \} $ and degree of idiosyncrasy of individual valuations $\beta \in \{ 0, 0.4, 0.8, 1, 2, 5\}$. When we consider results varying a single parameter, we group the results over the other parameters and provide reports varying only the parameter of interest. We report results for a relatively small consumption history $T=20$ with a produce space size $N=200$. Unless otherwise reported, our results are robust to varying $N$.

@@

\begin{figure}[t]
\includegraphics[width=\linewidth]{figures/gamma_consumption_dist_N_200T_20.pdf}\\
\includegraphics[width=\linewidth]{figures/gamma_consumption_dist_N_200T_20_partial.pdf}\\
\includegraphics[width=\linewidth]{figures/gamma_consumption_dist_N_200T_20_omni.pdf}\\
\caption{Effect of Risk Aversion on Local Consumption for No Recommendation (Top), Recommendation (Middle) and Oracle (Bottom)}
\label{fig:no_rec_risk_aversion}
\end{figure}

@@

\begin{figure}[ht]
\includegraphics[width=.9\linewidth]{figures/gamma_diversity_N_200_T_20}
\captionof{figure}{Risk Aversion and Diversity}\label{fig:risk_aversion_diversity}
\end{figure}

\begin{finding}\label{finding_welfare_gap}
The welfare gap between the no recommendation and the recommendation regimes and that between the no recommendation and the oracle regimes are decreasing in $\rho$.
\end{finding}
However, interestingly, the decrease in diversity is \textit{not} associated with a decline in welfare. In fact, welfare stays roughly the same in the oracle and recommendation cases. The welfare gap between the no recommendation case and the other two cases \textit{decreases} as the diversity gap increases.
\par
\begin{figure}[t]
\includegraphics[width=.9\linewidth]{figures/rho_diversity_N_200_T_20.pdf}\\
\includegraphics[width=.9\linewidth]{figures/rho_welfare_N_200_T_20.pdf}\\
\captionof{figure}{Welfare, Diversity, and Correlation}\label{fig:diversity_correlation}
\end{figure}

@@

This only happens since $\gamma = 0$ leads to users only caring about the expected value of the product. However, as we saw in Findings \ref{finding_local_consumption} and \ref{finding_diversity}, increasing $\gamma$ can lead to lower diversity and increasingly local consumption due to the fact that the degree of uncertainty now impacts users' choices. This weakens the negative relationship between diversity and welfare since both negative and positive experiences with a good reduce uncertainty about surrounding goods. This leads to the inverted-U shape found in Figure \ref{fig:diversity_welfare_ra} when $\gamma$ is relatively large (e.g $\gamma = 5$) though diversity and welfare are virtually uncorrelated in the data. In the recommendation and oracle regimes, under risk neutrality ($\gamma=0$), welfare and diversity and uncorrelated, while under risk aversion ($\gamma=5$), it is possible to observe in Figure \ref{fig:diversity_welfare_ra_partial} an actual positive relation between diversity and welfare as recommendations are able to reduce uncertainty and facilitate exploration of the product space.
\par
\begin{figure}[t]
\includegraphics[width=1.05\linewidth]{figures/diversity_welfare_rn.pdf}\\
\includegraphics[width=1.05\linewidth]{figures/diversity_welfare_ra.pdf}\\
\captionof{figure}{Diversity vs. Welfare, No-Recommendation, $\gamma = 0$ (Top), Diversity vs. Welfare, $\gamma = 5$ (Bottom)}\label{fig:diversity_welfare_ra}
\end{figure}
\begin{figure}[t]
\includegraphics[width=1.05\linewidth]{figures/diversity_welfare_rn_partial.pdf}\\
\includegraphics[width=1.05\linewidth]{figures/diversity_welfare_ra_partial.pdf}\\
\captionof{figure}{Diversity vs. Welfare, Recommendation, $\gamma = 0$ (Top), Diversity vs. Welfare, $\gamma = 5$ (Bottom)}\label{fig:diversity_welfare_ra_partial}
\end{figure}

@@

\begin{figure}[t]
\includegraphics[width=.8\linewidth]{figures/beta_homogeneity_N_200_T_20}
\captionof{figure}{Strength of Recommendation and Homogeneity}\label{fig:beta_homo}
\end{figure}
~

\begin{figure}[t]
\includegraphics[width=.8\linewidth]{figures/rho_homogeneity_N_200_T_20}
\captionof{figure}{Correlation and Homogeneity}\label{fig:cor_homo}
\end{figure}

@@

As a result, several alternative recommendation evaluation metrics, such as serendipity \cite{kotkov2016survey}, coverage \cite{ge2010beyond}, novelty \cite{vargas2011rank}, and many others have been proposed in order to develop recommender systems that produce recommendations that are more useful for users. We use our model to motivate an alternative recommendation evaluation metric that provides an alternative perspective on the serendipity metric proposed in the literature. Serendipitous recommendations are said to ``have the quality of being both unexpected and useful" \cite{maksai2015predicting} and we use this as a guide towards our definition.

@@

We have used the insights from our model to guide the evaluation and design of recommender systems. The key insight is that a critical component of designing good recommendations is understanding what user behavior would be without recommendation and then understanding how recommendation would change their consumption. Operationalizing this requires collecting information on user beliefs about the goods in the product space, which is not data that is traditionally collected for the development of recommender systems.
\par
